Approximation as Differentiation

### 2025-05-08 main/formal-structures.tex
Intro to Fam section.

@@

First we introduce the category of lattices and Galois connections, which will serve as our base category for
interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
of $g$, and $g$ the best approximation from above to an inverse of $f$.

>>>>

\subsection{A categorical framework for Galois slicing}

We will consider a total language where types denote sets. The goal is to interpret programs as Galois
connections relating approximations of their input to approximations of their output. Since the appropriate
lattice of approximations depends on the value being approximated, the semantics of every type $\tau$ must
have both a ``conventional'' component assigning $\tau$ to a set $X$, and an associated lattice of
approximations $\partial X(x)$ for every point $x \in X$. An expression $e$ of type $Y$ with a single free
variable of type $X$ must be interpreted as a morphism that combining both a function $f: X \to Y$ and a
family of Galois connections $\partial f(x): \partial X(x) \to \partial Y(f(x))$ for every $x \in X$.

@@

First we introduce the category of lattices and Galois connections, which will serve as our base category for
interpreting forward and backward Galois slicing. Galois connections are pairs of monotone functions $f: Y \to
X$ and $g: X \to Y$ between posets, where $f$ is the (pointwise) best approximation from below to an inverse
of $g$, and $g$ the best approximation from above to an inverse of $f$. This setup is a nice fit for the
bidirectional nature of (dynamic) program slicing, with $f$ capturing \emph{backward} slicing (producing the
least input slice for a given slice of the output), and $g$ capturing \emph{forward} slicing (producing the
greatest output slice for a given slice of the input).
